Detecting social information in a dense dataset of infants’ natural visual experience

### 2020-01-24 paper/cogsci/paper.Rmd
cogsci paper draft with analysis integrated

---
title: "A longitudinal investigation of the social information in natural infant visual experience"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract: >
    The faces and hands of infants' caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. 
    Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. 
    The present research aims to test the generality of these findings using the SAYcam dataset [@SAYcam], a longitudinal collection of over 1700 headcam videos collected from three children along a span of 6 to 32 months of age. 
    To do so, we validate the use of a modern convolutional neural network for pose detection (OpenPose) for the detection of people, faces, and hands to analyze these naturalistic infant egocentric videos. 
    We then apply this model to the entire dataset, analyzing the prevalence of faces across age, individuals, and activity contexts. 
    Overall, we find a higher prevalence of hands seen by infants than previously reported, considerably variability in the proportion of faces/hands seen across different locations (e.g., living room vs. kitchen), yet surprising consistency across both individual children.

@@

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

@@

# Introduction

Infants are confronted by a blooming, buzzing onslaught of stimuli (James, 1891) which they must learn to parse to make sense of the world around them. 
Yet infants do not embark on this learning process alone: infants are engaged in learning from their caregivers from early infancy. 
From as early as 3 months of age, young infants follow overt gaze shifts [@Gredeback2010], and even newborns prefer to look at faces with direct vs. averted gaze [@Farroni2002], despite their limited acuity [CITE].  
Faces are thus likely to be an important conduit of social information that scaffolds infants cognitive development. 
Given this importance, developmentalists have long hypothesized [CITE] that faces are prevalent in the visual experience of young infants. 
However, as even the viewpoint of a walking child is not easily predicted by our own adult intuitions (Clerkin, Hart, Rehg, Yu, & Smith, 2017; @Franchak2011; @Yoshida2008), researchers have begun to record the egocentric views collected from infants and toddlers wearing head-mounted cameras to test theories about the infant perspective.

A growing body of work now demonstrates that the viewpoints of very young infants--less than 4 months of age--are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @Sugden2014; @Jayaraman2018]. 
However, as infants mature, their perspective starts to capture views of hands paired with the objects they are acting on [@Fausey2016]. 
As infants learn to use their own hands to act on the world, they may focus on manual actions taken by their social partners. 
Furthermore, caregivers may start to use their hands more with communicative intent, directing infants attention with pointing and gestures to particular events or objects during play (Yu & Smith, 2013).

The present research aims to test the generality of these findings using the SAYcam dataset [@SAYcam], a longitudinal collection of over 1700 headcam videos collected from three children along a span of 6 to 32 months of age. 
In addition to its size and longitudinal nature, this dataset builds on those used in previous research in two key ways. 
First, recordings were from a variety of many different naturalistic contexts, encompassing infants’ viewpoints during both activities outside and inside the home. 
Second, the cameras used in this longitudinal study encompassed a much wider field of view than those typically used, allowing a more complete picture of the infant perspective.

However, with hundreds of hours of footage (>30M frames), this large dataset truly necessitates a shift to an automated annotation strategy. 
Indeed, annotation of the frames extracted from egocentric videos has been to be prohibitively time-consuming, meaning that many of the frames are not inspected. 
For example, @Fausey2016, collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand annotated (by four coders), totalling 103,383 frames (per coder)--an impressive number of annotations but nonetheless only 0.67% of the collected footage. 
To address this challenge, we first validate the use of a modern computer vision model [@Cao2018openpose] to automatically detect the presence of hands and faces from the infant egocentric viewpoint. 
In particular, we focus on OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints (135 in total) that operates well on scenes including multiple people even if they are partially-occluded (see Figure 1).  
We then apply these methods at scale to the larger dataset, allowing us to analyze the proportion of faces and hands observed by each child across age and activity context.

In the following paper, we first describe the dataset and the pose model used in the following analyses and validate the use of this model for extracting our key descriptive variables by comparing to a human-annotated gold set of 24,000 frames. 
Next, we analyze key descriptive variables, including those that have been previously reported to vary across age, including the relative proportions of faces vs. hands and the sizes of the faces, over the entire dataset (30M frames), finding a greater prevalence of hands than has been previously reported.  
We then investigate sources of variability in our more naturalistic dataset that may explain these differences, including a diversity of activity contexts as well as a larger field of view captured by our cameras.

@@

## Dataset

Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase 109 degrees horizontal x 70 degrees vertical. 
Children wore headcams at least twice weekly, for approximately one hour per recording session. 
One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family's discretion. 
At the time of the recording, all three children were in single-child households. 
Videos[^1] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset.
We analyze 1,636 videos, with a total duration of XX hours (XX million frames).

[^1]: All videos are available at https://nyu.databrary.org/volume/564

## Part 1: How well can we capture social information using computer vision?

### Computer vision model
To automatically annotate the millions of frames in SAYcam, we use OpenPose [@Cao2018openpose; @Simon2017hand], a computer vision model optimized for jointly detecting human face, body, hand, and foot keypoints (135 in total) that operates well on scenes including multiple people even if they are partially-occluded. 
This convolutional neural network (CNN)-based pose detector[^2] provided the locations of 18 body parts (ears, nose, wrists, etc.). 
The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. 
The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of outputs of the face and hand detection modules.

@@

### Manual annotation strategy
To test the validity of OpenPose's hand and face detections, we compared the accuracy of these detections relative to human annotations of 24,000 frames selected uniformly at random from the 1,636 videos of two children (S and A).
Frames were jointly annotated for the presence of faces and hands. 
These randomly sampled frames covered XX of the videos present in the dataset.  
A second set of coders recruited via AMT (Amazon Mechanical Turk) additionally annotated XX frames; agreement with the primary coder was XX%.

### Detection accuracy for faces and hands
Overall, we found that face detections were slightly more accurate than hand detections,

Precision and recall (F-score) variation across child/age for faces

Describe possible sources of variation that decrease scores for:
-Faces: weird viewpoints, occluded/side viewpoint, faces in books
-Hands: children's own hands, hands in books, side viewpoints

Describe additional child vs. hand annotation; P/R/F variation across child vs. adult hands (better for adult hands, still OK for child hands)

## Part 2: Access to social information across age

Next, we analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child. 
Data from videos were binned according to the age of the child (in weeks) (see Figure XX). 
First, we saw that the proportion of faces in view showed a moderate decrease across this age range, both when analyzing the random 24K frames as well as the entire dataset. 
While these trends appear somewhat different than those observed in @Fausey2016, note that @Fausey2016 included data from very young infants (starting at 2 months while), while here the youngest videos coming from S and A around 6 and 9 months of age, respectively. 
Similarly, our age range extends 8 months later than those infants in @Fausey2016, throughout a portion of third year of life.

However, the most striking result from the dataset is a much greater proportion of hands in view than have previously been reported. 
We found this to be true across all ages, in both children, and regardless of whether we analyzed human annotations (on the 24K random subset) or the entire dataset.   
One reason this could be the case is the much larger field of view that was captured by the cameras used in this study: unlike previous studies, our cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible.  
Thus, the field of view (FOV) of the fisheye lens used in @SAYcam was much wider (109 degrees horizontal x 70 degrees vertical) than the FOV of the lens used in @Fausey2016 (69 x 41 degrees).
This larger field of view may have allowed the SAYcam cameras to capture not only the presence of a social partner’s hands interacting with objects, but also the children’s own hands, leading to more frequent hand detections.

To assess this possibility, we obtained annotations for a subsample of XX frames in which a hand was detected in the random gold set; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult hands. 
Overall, we found that 34% of the hands detected by OpenPose in the random 24K sample were of children’s own hands (compared to 8% reported in @Fausey2016) suggesting that this difference in field of view did contribute substantially to the higher rate of detections in the frames. 
Heatmaps of the bounding boxes obtained from these annotations can be seen in Figure XX, showing that children’s hands tended to appear in the lower half of the frames.

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, middle portion of the frame comparable to the field of view used in @Fausey2016 (XX vs XX degrees).
To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all detections that occurred in the top of the frame.  
Overall, we found X, suggesting Y.

Intriguingly, we observed an unexpected trend in which the relative proportion of faces vs. hands increased during the third year of life; in particular, this seemed...(save for discussion?)

### Variability by Location 
Next we examine variation in the presence of hands and faces across different locations. 
Of the XX videos, the content of 1,829 have been manually annotated for filming location, activites taking place, and visible objects (see @SAYcam).
XX of the videos were filmed in a single location, representing XX% of the dataset and roughly XX million frames (see @SAYcam). 
To give a sense of the contexts the children experienced, the most frequent filming locations were the living room (339 videos), bedroom (182), kitchen (150), outside on property (129), child's bedroom (81), deck/porch (73), hallway (70), and off property (57). 
Filming only took place twice in the dining room (see Figure XX). 
Activities varied predictability by these locations: for example, eating and XX occured in the kitchen, whereas playtime was the dominant activity in the living room (see Figure XX). 
Bria: analyze activity

Not this, but point out that eating was not frequent:
The most frequent activities were sitting (410), playing (375), being held (352), and standing (297). 
Eating was the 11th most-frequent activity (117 videos).

(goldset, full dataset)

# Discussion

We demonstrate the feasibility of using modern computer vision models to vastly increase the efficiency of processing egocentric headcam footage, allowing us to annotate 100% of even very large datasets for the presence and size of people, hands, and faces.
We validate the use of this model by comparing to a human-annotated gold set, and find X.

Analyzing this dataset has yielded a better understanding of infants' evolving access and attention to social information.

Need to think about the child's viewpoint relative to actual FOV as well as attention

Fausey 2016: 103,383 images; 
Here: 30,000,000 frames; 300 fold increase in data

>>>>

---
title: "Detecting social information in a dense dataset of infants’ natural visual experience"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract: >
      The faces and hands of infants' caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validate the use of a modern convolutional neural network for pose detection (OpenPose) for the detection of faces and hands. We then apply this model to the entire dataset, finding a decrease in the proportion of faces vs. hands across age up to the end of the second year of life [@Fausey2016]. In addition, we find considerable variability in the proportion of faces/hands viewed by children in different locations (e.g., living room vs. kitchen), demonstrating that variability across activity contexts may be a driving factor in the amount of social information that infants’ experience.

@@

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---
\newcommand{\wrapmf}[1]{#1}

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)

```

@@

# Introduction
Infants are confronted by a blooming, buzzing onslaught of stimuli [@James1981] which they must learn to parse to make sense of the world around them. Yet they do not embark on this learning process alone: From as early as 3 months of age, young infants follow overt gaze shifts [@Gredeback2010], and even newborns prefer to look at faces with direct vs. averted gaze [@Farroni2002], despite their limited acuity. As faces are likely to be an important conduit of social information that scaffolds cognitive development, cognitive scientists have long hypothesized that faces are prevalent in the visual experience of young infants.

Yet until recently most hypotheses about infants’ visual experience have gone untested. Though parents and scientists alike have strong intuitions about what infant’s see, even the viewpoint of a walking child is not easily predicted by these intuitions (Clerkin, Hart, Rehg, Yu, & Smith, 2017; Franchak et al., 2011; Yoshida & Smith, 2008). By equipping infants and toddlers with head-mounted cameras, researchers have begun to document the infant egocentric perspective on the world. Using these methods, a growing body of work now demonstrates that the viewpoints of very young infants (less than 4 months of age) are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @Sugden2014; @Jayaraman2018].

As infants motor and cognitive abilities mature beyond these early months, their perspective on the world becomes even more varied. Crawlers see fewer faces and hands than do walking children [@Sanchez2018; @Franchak papers] as well as different view of objects [@Soska2010].  Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, as their perspective starts to capture views of hands paired with the objects they are acting on [@Fausey2016]. In turn, caregivers may start to use their hands with more communicative intent, directing infants attention with pointing and gestures to different events or objects during play [@YuSmith2013].

Here, we analyze the social information present in the infant visual perspective by analyzing a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of contexts, encompassing infants’ viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant’s experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth).  Second, the head-mounted cameras used in the SAYcam dataset captured a substantially larger field of view than those typically used, allowing a more complete picture of the infant perspective. Head-mounted cameras with a more restricted field of view do still capture where infants are foveating most of the time [@YoshidaSmith2008, @SmithFausey-Review] but may still fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

With hundreds of hours of footage (>29M frames), this large dataset necessitates a shift to an automated annotation strategy. Indeed, annotation of the frames extracted from egocentric videos has been prohibitively time-consuming, meaning that many frames are typically not inspected, even in the most comprehensive studies. For example, @Fausey2016 collected a total of 143 hours of head-mounted camera footage (15.5 million frames), of which one frame every five seconds was hand annotated (by four coders), totalling 103,383 frames (per coder)—an impressive number of annotations but nonetheless only 0.67% of the collected footage. To address this challenge, we use a modern computer vision model of pose detection to automatically detect the presence of hands and faces from the infant egocentric viewpoint. Specifically, we use OpenPose [@Cao2018openpose], a model optimized for jointly detecting human face, body, hand, and foot keypoints that operates well on scenes including multiple people, even if they are partially-occluded (see Figure 1).

In the following paper, we first describe the dataset and validate the use of this model by comparing face and hand detections to a human-annotated set of 24,000 frames. Next, we report how the proportion of faces and hands changes across age. We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the present cameras as well as a diversity of locations during which videos were recorded.

```{r}
## Read in data 
### Load preprocessed and filtered detections
load('../../data/openpose_detections_filtered_threekids.RData') #d_all
load('../../data/openpose_detections_filtered_cropped_allthreekids.RData') # d_cropped
```

```{r}
### Choose bins across which to analyze data
bin_size = 7 # days
min_age = min(d_all$age_days)
max_age = max(d_all$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)
```

```{r}
## Make summary stats for cropped and uncropped detedctions
summary_by_age <- d_all %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  group_by(age_day_bin, child_id) %>%
  summarize(num_detect = length(face_openpose), prop_faces = sum(face_openpose) / num_detect,  prop_hands = sum(hand_openpose) / num_detect)

summary_by_age_cropped <- d_cropped %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  group_by(age_day_bin,child_id) %>%
  summarize(num_detect = length(face_openpose), prop_faces = sum(face_openpose) / num_detect,  prop_hands = sum(hand_detected_cropped) / num_detect) 
```

```{r} 
### How large is the dataset we're analyzing?
num_frames = length(d_all$frame)

fps=30
num_seconds = length(d_all$frame)/fps
num_videos = length(unique(d_all$vid_name))
num_minutes = num_seconds/60
num_hours = num_minutes/60
num_days = num_hours/24 
```

@@

## Dataset
The dataset is described in detail in @SAYcam; we summarize these details here. Children wore Veho Muvi miniature cameras mounted on a custom camping headlamp harness (“headcams”) at least twice weekly, for approximately one hour per recording session. One weekly session was on the same day each week at a roughly constant time of day, while the other(s) were chosen arbitrarily at the participating family’s discretion. At the time of the recording, all three children were in single-child households.  Videos captured by the headcam were 640x480 pixels, and a fisheye lens was attached to the camera to increase the field of view to approximately 109 degrees horizontal x 70 degrees vertical. Videos[^1] with technical errors or that were not taken from the egocentric perspective were excluded from the dataset. While data collection for the third child (Y) is still ongoing, here we analyze `r num_videos` videos, with a total duration of `r num_hours` hours (>29 million frames).
[^1]: All videos are available at https://nyu.databrary.org/volume/564

```{r examples, include = T, fig.env = "figure*", fig.pos = "h", fig.align='center', fig.width=8, fig.height=3.5, set.cap.width=T, num.cols.cap=2, fig.align = "center", fig.cap = "Example frames taken from the dataset, illustrating variability in the infant perspective across different locations. OpenPose detections are shown overlaid on these images."}
examples <- png::readPNG("figs/modifiedOPfigure_v3.png")
grid::grid.raster(examples)

```

## Assessing automated social annotations

### Computer vision model
To automatically annotate the millions of frames in SAYcam, we use OpenPose [@Cao2018openpose; @Simon2017hand]. This pose detector[^2] provided the locations of 18 body parts (ears, nose, wrists, etc.). The system uses a convolutional neural network for initial anatomical detection and subsequently applies part affinity fields for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. Thus, while we only make use of the outputs of the face and hand detections, the entire set of pose information from an individual is used to determine the presence of a face/hand, making the process much more robust to occlusion than methods optimized to detect only faces or hands. Specifically, we tagged frames with detected nose keypoints as having a face, and frames with a wrist keypoint were tagged as having a hand.

@@

### Manual annotation strategy
To test the validity of OpenPose's hand and face detections, we compared the accuracy of these detections relative to human annotations of 24,000 frames selected uniformly at random from videos of two children (S and A). Frames were jointly annotated for the presence of faces and hands.  A second set of coders recruited via AMT (Amazon Mechanical Turk) additionally annotated 3150 frames; agreement with the primary coder was >95%.

### Detection accuracy for faces and hands

```{r}
### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
right_side_up_file = "../../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos
flipped_videos <- right_side_up %>%
  filter(right_side_up==1)

## load goal sample annotations
load('../../data/ketan_gold_sample.RData')

## preprocess them
 ketan_gold_out <- ketan_gold %>%
   rowwise() %>%
   mutate(vid_name = str_split_fixed(vid_name,".mp4",2)[,1])%>%
   mutate(vid_name_short = str_split(vid_name,"[.]")[[1]][1]) %>%
   filter(!vid_name_short %in% flipped_videos$video) %>%
   mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
   rename(face_present_ketan = face_present, hand_present_ketan = hand_present)

# Function to evaluate detectors
evaluate_detector <- function(truth, detection) {
  if (truth == TRUE) {
    if (truth == detection) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (truth == detection) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}

# function to return prfs
 return_prf_short = function(eval){
  tp=sum(eval == "TP")
  fp=sum(eval == "FP")
  fn=sum(eval == "FN")
  p = tp / (tp + fp)
  r = tp / (tp + fn)
  f=( 2 * p * r )/ (p + r)
  return(c(p,r,f))
 }

 ## evaluate detectors
 gold_sample <- ketan_gold_out %>%
   mutate(face_eval_ketan = evaluate_detector(face_present_ketan, face_openpose), hand_eval_ketan = evaluate_detector(hand_present_ketan, hand_openpose))

 ## output prfs
 face_performance = return_prf_short(gold_sample$face_eval_ketan)
 hand_performance = return_prf_short(gold_sample$hand_eval_ketan)

```
As has been observed in other studies on automated annotation of headcam data [e.g. @frank2013, @REF-chenyu-hand-paper], detection tasks that are easy in third-person video can be quite challenging in egocentric videos, due to difficult angles and sizes as well as substantial occlusion. For example, the infant perspective often contains non-canonical viewpoints of faces (e.g., looking up at a caregiver's chin) as well as partially-occluded or oblique viewpoints of both faces and hands.

To evaluate OpenPose's performance, we compared its detections to the manually-annotated gold set of frames, calculating precision (hits / hits + false alarms), recall (hits / hits + misses), and f-score (the harmonic mean of precision and recall).  In our data, for faces, the F-score was `r round(face_performance[3],2)`, with a precision of `r round(face_performance[1],2)` and recall of `r round(face_performance[2],2)`. For hands, the F-score was `r round(hand_performance[3],2)`, with a precision of `r round(hand_performance[1],2)` and recall of `r round(hand_performance[2],2)`. While face and hand detections showed moderately good precision, face detections were overall slightly more accurate than hand detections. Hand detections suffered from fairly low recall, indicating that OpenPose likely underestimated the proportion of hands in the dataset. We suspect that this is in part because children’s own hands were often in view of the camera and unconnected to a pose; this factor is explored in detail in the following analysis section.

We also examined whether the precision, recall, and F-score for hands and faces varied per child as a function of age to see if there was any bias in the OpenPose detections related to age, and did not find substantial variation. Nonetheless, while OpenPose was trained on photographs from the adult perspective, this model still generalized relatively well to the egocentric infant viewpoint with no fine-tuning or post-processing of the detections.

## Access to social information 
### Changes across age

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands that were in view for each child. Data from videos were binned according to the age of the child (in weeks) (see Figure 2). First, when examining the younger age ranges (< 24 months) we saw that the proportion of faces in view showed a moderate decrease across this age range, both when analyzing the random 24K frames as well as the entire dataset (see Figure 2A).

While these trends appear somewhat different than those observed in @Fausey2016, note that @Fausey2016 included data from very young infants (from 1-4 months of age), while here the youngest videos coming from S and A around 6 and 9 months of age, respectively.  Similarly, our age range extends 8 months later than those infants in @Fausey2016, throughout a portion of third year of life.

```{r FaceVsHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces vs. hands in the entire dataset (A) and when hand detections are restricted to the upper 60 percent of the field of view (B). Data from each child (A, S) are plotted separately and binned by each week that the videos were filmed; data is scaled by the number of frames in that age range."}

faces <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  ylab('Proportion Faces') + 
  # theme_few(base_size=16) +
  xlab('Age (months)') +
  ylim(0,.5) +
  theme(legend.position="none") +
  ggtitle('(A).') +
  scale_colour_brewer(palette = "Dark2")

hands <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Hands') + 
  xlab('Age (months)') +
  ylim(0,.5)+
  theme(legend.position="none") +
  ggtitle('(B).') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand <- ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Faces - Hands') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('(C).') +
  scale_colour_brewer(palette = "Dark2")

face_v_hand_cropped <- ggplot(summary_by_age_cropped, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10) + 
  # theme_few(base_size=16) +
  ylab('Proportion Faces - Hands') + 
  xlab('Age (months)') +
  ylim(-.3, .3) +
  theme(legend.position = "none") +
  ggtitle('(D).')  +
  scale_colour_brewer(palette = "Dark2")

ggarrange(faces, hands, face_v_hand, face_v_hand_cropped, nrow=1)
```

However, the most striking result is a much greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in both children, and regardless of whether we analyzed human annotations (on the 24K random subset) or OpenPose annotations on entire dataset. This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this is still an underestimate of the proportion of hands in view. One reason this could be the case is the much larger field of view that was captured by the cameras used in this study: unlike previous studies, our cameras were outfitted with a fish-eye lens in an attempt to capture as much of the children’s field of view as possible. Thus, the field of view (FOV) of the fisheye lens used was larger (109 degrees horizontal x 70 degrees vertical) than the FOV of the lens used in many previous studies; for example, in @Fausey2016 the FOV was 69 x 41 degrees. This larger field of view may have allowed the SAYcam cameras to capture not only the presence of a social partner’s hands interacting with objects, but also the children’s own hands, leading to more frequent hand detections.

```{r}
turk_annotations_seg = "../../data/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

child_hands <- g_seg %>%
  filter(label=="Child hand")

adult_hands <- g_seg %>%
  filter(label=="Adult hand")

##
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  # ggtitle('(A). Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=12) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) 
  # ggtitle('(B). Adult hand density') 
```

```{r density, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=4, fig.height=2, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in teh 24K random gold set; each dot represents the center of a bounding box made by an adult participant."}
ggarrange(child_hand_plot, adult_hand_plot, nrow=1)
```

To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected in the random gold set; participants (recruited via AMT) were asked to draw bounding boxes around children’s hands and adult’s hands. Overall, we found that 34% of the hands in the random 24K sample were of children’s own hands (compared to 8% reported in @Fausey2016) suggesting that this difference in field of view did contribute substantially to the higher rate of hand detections in the frames. Heatmaps of the bounding boxes obtained from these annotations can be seen in Figure 3, showing that children’s hands tended to appear in the lower half of the frames.

We thus re-analyzed the entire dataset while restricting our analysis to a smaller, upper portion of the frame comparable to the field of view used in @Fausey2016. To do so, we excluded hand detections that occurred in the bottom 40% of the frame, while retaining all hand detections that occurred in the top 60% of the frame. This coarse cropping of decreased the proportion of hand detections from `r mean(d_all$hand_openpose)`percent to `r mean(d_cropped$hand_detected_cropped)` percent.

With this modified field of view, we now observed a more substantive decrease across age in the proportions of faces relative to hands. To quantify this trend, we fit a generalized linear mixed model to the binned detection rates from this smaller field of field to estimate changes in the proportion of faces vs. hands over age. To better approximate @Fausey2016, we also restricted our analysis to the age range seen in @Fausey2016, excluding videos when children were over 24 months of age, and indeed found that the proportion of faces vs. hands declined modestly with age (see Table 1).

However, when we considered the entire age range, we also observed an unexpected trend in which the relative proportion of faces vs. hands increased again after 24 months of life; this trend was stronger in the full dataset. While we did not have any a priori hypotheses about this age range, we suspect that this could have to do with an increase in children’s independence from their caregiver as they master walking around and interacting with objects in the world on their own.

### Variability by Location 
```{r}
load('../../data/child_adult_hand_annotations.RData')
load('../../data/saycam_metadata.RData')

num_videos_annotated = sum(!is.na(meta$Location))
videos_included = sum(!is.na(meta$Location) & meta$count_locations==1)

# percentage_of_dataset = sum(faces_vs_hands_to_plot$num_detect)/length(d_all$frame)
# num_frames = sum(faces_by_location_to_plot$num_detect)
```

How does variability across different contexts influence the social information in the infant view? 
Intuitively, some activities in different contexts may be characterized by a much higher proportion of faces (e.g., diaper changes in bedrooms) than others (e.g., playtime in the living room). We thus next examined variation in presence of hands and faces across different locations. Of the `r num_videos` videos, `r num_videos_annotated` were annotated [@SAYcam] for the location o videos were filmed in. `r videos_included` were filmed in single location, representing 17 percent of the dataset and over 5 million frames (see @SAYcam). Activities varied somewhat predictability by these contexts: for example, eating tended to occur in the kitchen, whereas playtime was the dominant activity in the living room. Overall, we found that the proportion of faces vs. hands varied across filming locations, and, to some extent, across children.  For example, while both A and S saw a relatively similar proportion of faces vs. hands in the bedroom, they saw quite different amounts of faces vs. hands in kitchen.

```{r}
d_by_location_cropped <- d_cropped %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  left_join(meta) %>%
  filter(!is.na(Location)) %>%
  filter(count_locations==1) 
# 
# d_by_location_cropped$Location = droplevels(d_by_location_cropped$Location)
# levels(d_by_location_cropped$Location)

face_hand_by_location_cropped <- d_by_location_cropped %>%
  group_by(Location,child_id,age_day_bin) %>%
  summarize(prop_faces = mean(face_openpose), prop_hands = mean(hand_detected_cropped), num_detect = length(hand_detected_cropped))  %>%
  filter(num_detect > 12000) %>% ## 18000 frames = 10 minutes of video at 30 fps %>%
  mutate(faces_vs_hands = prop_faces - prop_hands)

hands_to_plot_cropped <- face_hand_by_location_cropped %>%
  multi_boot_standard(col="prop_hands") %>%
  ungroup %>%
  left_join(face_hand_by_location_cropped) %>%
  filter(!is.na(ci_lower)) %>% # if we didn't have enough points to make a CI, filter
  mutate(Location = fct_reorder(Location, mean))

faces_to_plot_cropped <- face_hand_by_location_cropped %>%
  multi_boot_standard(col="prop_faces") %>%
  ungroup %>%
  left_join(face_hand_by_location_cropped) %>%
  filter(!is.na(ci_lower)) %>% # if we didn't have enough points to make a CI, filter
  mutate(Location = fct_reorder(Location, mean)) 
```

```{r DetByLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=4, fig.cap = "Proportion of faces (A),  hands (B), and faces vs. hands  (C) by location in which egocentric videos were filmed; data from individual children are plotted separately (location annotations were not available for Y). Data are only taken from the top 60 percent of the frames so as to minimize the contribution of children’s own hands. Each dot represents data from a week in which videos were filmed and are scaled by the number of frames (i.e., amount of video) in that datapoint."}

faces_loc = ggplot(faces_to_plot_cropped, aes(x = Location, y = mean, col=child_id)) + 
  geom_point(aes(x=Location, y=prop_faces, size=num_detect), alpha=.2, position = position_dodge(width=.3)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width=.3)) +
  # ylim(-.4,.4) +
  coord_flip() + 
  ylab('Proportion Faces')+
  xlab('')+
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")

hands_loc = ggplot(hands_to_plot_cropped, aes(x = Location, y = mean, col=child_id)) + 
  geom_point(aes(x=Location, y=prop_hands, size=num_detect), alpha=.2, position = position_dodge(width=.3)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width=.3)) +
  # ylim(-.4,.4) +
  coord_flip() + 
  xlab('')+
  ylab('Proportion Hands') +
  scale_colour_brewer(palette = "Dark2") +
  theme(legend.position = "none")

ggarrange(faces_loc, hands_loc, nrow=2)
```
# General Discussion
